---
title: "DyNet paper code: Prelims & VAR dynamics"
author:
    name: Joshua Cape
    affiliation: University of Pittsburgh
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

## Preliminaries
```{r message=FALSE}

    rm(list = ls())

    knitr::opts_chunk$set(comment = NA)

    source("code_dynet_preamble.R")

```

## Initialization
```{r}

    set.seed(12345)
    lp.dim <- 2;
    lp.num <- 200;
    
#######################
    
    lpvs0 <- diag(lp.dim)
    
    lpvs <- rbind(lpvs0,
                  t(apply(matrix(rnorm((lp.num - dim(lpvs0)[1]) * lp.dim, mean=0, sd=1),
                                 nrow=(lp.num - dim(lpvs0)[1]), ncol=lp.dim),
                             1,
                             function(x) (x/norm(x, "2"))*runif(1, min=0, max=1)^(1/lp.dim))))
    
    XXT <- lpvs %*% t(lpvs)
    Prob <- abs(XXT)
    A <- matrix(runif(lp.num*lp.num) <= Prob, nrow=lp.num)

    matrix.A <- sym(A) * sign(XXT)
    
    my.eigen <- partial_eigen(matrix.A, lp.dim, symmetric=TRUE)
    my.scaled <- my.eigen$vectors %*% abs(diag(my.eigen$values))^(1/2)

    my.candidate.svd <- svd(my.scaled[1:lp.dim,])
    my.candidate.orthog <- my.candidate.svd$u %*% t(my.candidate.svd$v)
    my.supervised.orthog <- my.candidate.orthog
   
    theory.svd <- svd(t(lpvs) %*% my.scaled)
    theory.orthog <- theory.svd$u %*% t(theory.svd$v)
     
```


## Introductory plots: visualizing latent vectors and supervised embedding
```{r}

        dat.circle <- circleFun(c(0,0), 2, npoints = 100)
    
        dat.circle.ttinf <- circleFun(c(0,0), 2*ttinf(my.scaled %*% t(theory.orthog) - lpvs), npoints = 100)
    
        temp.df1 <- as.data.frame(lpvs)
        colnames(temp.df1) <- c("x","y")
        
        temp.df2 <- as.data.frame(my.scaled %*% t(my.supervised.orthog))
        colnames(temp.df2) <- c("x","y")
        
        temp.df3 <- as.data.frame(my.scaled)
        colnames(temp.df3) <- c("x","y")

    temp.g1 <- ggplot(dat.circle, aes(x,y)) + geom_path() + coord_fixed(ratio=1) +
      geom_point(data=temp.df1, aes(x, y), pch=0)
    
    temp.g2 <- ggplot(dat.circle, aes(x,y)) + geom_path() + coord_fixed(ratio=1) +
      geom_point(data=temp.df2, aes(x, y))
    
    temp.g3 <- ggplot(data=dat.circle.ttinf, aes(x,y)) +
      xlim(-0.4, 0.4) + ylim(-0.4, 0.4) + coord_fixed(ratio=1) +
      
    geom_point(data=dat.circle.ttinf, aes(x,y), lwd=1/10) +
      geom_point(data=(temp.df1-temp.df2), aes(x, y), pch=18) + 
        geom_point(data=(temp.df1-temp.df3), aes(x, y), pch=3)

```

```{r}

  temp.g1 + temp.g2 + temp.g3

```


## Example: single node dynamics among few dynamic nodes
```{r}
    
    dynamic.node.num <- 10
    lp.num <- 200
    t.max <- 50
    
    # specify rotation matrices via hyperparameter thetas
    hyper.thetas <- as.list(c(rep(0,lp.dim),
                                (2*pi)*(1/dynamic.node.num)*seq(1,dynamic.node.num)))
    
    trans1 <- lapply(hyper.thetas, rot_mtx)
    
    lpvec.list <- list()
        lpvec.list[[1]] <- lpvs
    
    for(t in 2:t.max){
      lpvec.list[[t]] <- matrix(0, nrow=lp.num, ncol=lp.dim)
      for(lp in 1:lp.dim)
        lpvec.list[[t]][lp,] <- lpvec.list[[t-1]][lp,]
      for(lp in (lp.dim+1):(lp.dim + dynamic.node.num))
      {
        lpvec.list[[t]][lp,] <- lpvec.list[[1]][lp,] %*% matrix.power(t(trans1[[lp]]), t-1)
        }
      lpvec.list[[t]][(lp.dim + dynamic.node.num + 1):lp.num,] <- t(apply(
        matrix(rnorm((lp.num-lp.dim-dynamic.node.num)*lp.dim, mean=0, sd=1),
               nrow=(lp.num-lp.dim-dynamic.node.num),
               ncol=lp.dim),
        1,
        function(x) (x/norm(x, "2"))*runif(1, min=0, max=1)^(1/lp.dim)))
    }
    
    
    adj.list <- list()
    raw.scaled.list <- list()
    svsd.orthog.list <- list()
    
    Xhat.list <- list()
        Xhat.list[[1]] <- my.scaled %*% t(my.supervised.orthog)
    
    
    for(t in 1:t.max){
      
      A <- matrix(runif(lp.num*lp.num) <= abs(lpvec.list[[t]] %*% t(lpvec.list[[t]])), nrow=lp.num)
      adj.list[[t]] <- sym(A) * sign(lpvec.list[[t]] %*% t(lpvec.list[[t]]))
    
      my.eigen <- partial_eigen(adj.list[[t]], lp.dim, symmetric=TRUE)
      raw.scaled.list[[t]] <- my.eigen$vectors %*% abs(diag(my.eigen$values))^(1/2)
    
      my.candidate.svd <- svd(raw.scaled.list[[t]][1:lp.dim,])
      my.candidate.orthog <- my.candidate.svd$u %*% t(my.candidate.svd$v)
      svsd.orthog.list[[t]] <- my.candidate.orthog
    
      Xhat.list[[t]] <- raw.scaled.list[[t]] %*% t(svsd.orthog.list[[t]])
      
    }

```


### Single node dynamics - three optional plots
```{r}

which.node <- 3 # NB: first two are seed nodes

temp.mtx <- matrix(0, nrow=t.max, ncol=lp.dim)

##########
    
    for(t in 1:t.max)
      temp.mtx[t,] <- Xhat.list[[t]][which.node,]
    
    temp.df <- data.frame(cbind(temp.mtx, c(1:t.max), rev((1:t.max)/(10000*t.max))))
    colnames(temp.df) <- c("x","y","label", "size")
    
    dat.circle <- circleFun(c(0,0), 2, npoints = 100)
    
    p.scatter <- ggplot(dat.circle, aes(x,y)) + geom_path() + coord_fixed(ratio=1) +
       geom_point(data=temp.df,aes(x, y)) # + geom_text(data=temp.df,aes(label=label, size=size))
    
    p.index <- ggplot(temp.df, aes(x,y,label=label)) + geom_text(aes(label=label, size=size)) + theme(legend.position="none") + xlim(-1,1) + ylim(-1,1) + coord_fixed(ratio=1)
    
    p.traj <- ggplot(temp.df) +
      geom_line(aes(x=1:t.max, y=x), linetype="dashed") + geom_point(aes(x=1:t.max, y=x), shape=16) +
      geom_line(aes(x=1:t.max, y=y), linetype="dotted") + geom_point(aes(x=1:t.max, y=y), shape=17) + xlab("time") + ylab("x and y values") + coord_fixed(ratio=t.max/4)
    
```

```{r}

# Optional plot of theta_1 dynamics
#   p.scatter + p.index + p.traj

```



## Estimates via OLS, e.g., see Luetkepohl (2005)
```{r}

    Y.mtx <- t(temp.mtx)[,2:t.max]
    
    Z.mtx <- rbind(rep(1,(t.max-1)),
                   t(temp.mtx)[,1:(t.max-1)])
    
    # naive OLS estimate              
    Bhat.mtx <- Y.mtx %*% t(Z.mtx) %*% solve(Z.mtx %*% t(Z.mtx))
    
    # orthog projection estimate
    temp.decomp <- svd(Bhat.mtx[,2:3])
    temp.orthog <- temp.decomp$u %*% t(temp.decomp$v)
    
    # performance comparison
    target.orthog <- rot_mtx(hyper.thetas[[which.node]])
    
    mtx_print(Bhat.mtx[,1,drop=FALSE]) # OLS mu_Hat
    mtx_print(Bhat.mtx[,-1]) # OLS R_Hat
    mtx_print(target.orthog) # true R
    mtx_print(temp.orthog) # ORTH R_Hat

    norm(Bhat.mtx[,-1] - target.orthog, "F")/norm(target.orthog, "F")
    norm(temp.orthog - target.orthog, "F")/norm(target.orthog, "F")
```


# Re-run estimates under VAR(2)
```{r}
  
  Y.mtx <- t(temp.mtx)[,3:t.max]
  
  Z.mtx <- rbind(rep(1,(t.max-2)),
                 t(temp.mtx)[,2:(t.max-1)],
                 t(temp.mtx)[,1:(t.max-2)])
                 
  Bhat.mtx <- Y.mtx %*% t(Z.mtx) %*% solve(Z.mtx %*% t(Z.mtx))
  
  temp.decomp <- svd(Bhat.mtx[,2:3])
  temp.orthog <- temp.decomp$u %*% t(temp.decomp$v)
  
  target.orthog <- rot_mtx(hyper.thetas[[which.node]])
  
    mtx_print(Bhat.mtx[,1,drop=FALSE])
    #mtx_print(Bhat.mtx[,2:3])
    mtx_print(temp.orthog)
    #mtx_print(target.orthog)
    
  #norm(Bhat.mtx[,2:3] - target.orthog, "F")/norm(target.orthog, "F")
  norm(temp.orthog - target.orthog, "F")/norm(target.orthog, "F")
  
```

```{r}
  
  temp.decomp <- svd(Bhat.mtx[,4:5])
  temp.orthog <- temp.decomp$u %*% t(temp.decomp$v)
  
  target.orthog <- rot_mtx(hyper.thetas[[which.node]]);
  
    #mtx_print(Bhat.mtx[,1,drop=FALSE])
    #mtx_print(Bhat.mtx[,4:5])
    mtx_print(temp.orthog)
    #mtx_print(target.orthog %*% target.orthog)
    
  #norm(Bhat.mtx[,4:5] - target.orthog %*% target.orthog, "F")/norm(target.orthog, "F")
  norm(temp.orthog - target.orthog %*% target.orthog, "F")/norm(target.orthog, "F")

```


## Example: all non-seed nodes evolve according to individual dynamics   
```{r}
  
  t.max <- 400
  dynamic.node.num <- lp.num-lp.dim
  
  # specify rotation matrices via hyperparameter thetas
  hyper.thetas <- as.list(c(rep(0,lp.dim),
                              (2*pi)*(1/dynamic.node.num)*seq(1,dynamic.node.num)))
  
  trans1 <- lapply(hyper.thetas, rot_mtx)
  
  lpvec.list <- list()
      lpvec.list[[1]] <- lpvs
  
  for(t in 2:t.max){
    lpvec.list[[t]] <- matrix(0, nrow=lp.num, ncol=lp.dim)
    for(lp in 1:lp.dim)
      lpvec.list[[t]][lp,] <- lpvec.list[[t-1]][lp,]
    for(lp in (lp.dim+1):(lp.dim + dynamic.node.num))
    {
      lpvec.list[[t]][lp,] <- lpvec.list[[1]][lp,] %*% matrix.power(t(trans1[[lp]]), t-1)
      }
  }
  
  
  adj.list <- list()
  raw.scaled.list <- list()
  svsd.orthog.list <- list()
  
  Xhat.list <- list()
      Xhat.list[[1]] <- my.scaled %*% t(my.supervised.orthog)
  
  
  for(t in 1:t.max){
    
    A <- matrix(runif(lp.num*lp.num) <= abs(lpvec.list[[t]] %*% t(lpvec.list[[t]])), nrow=lp.num)
    adj.list[[t]] <- sym(A) * sign(lpvec.list[[t]] %*% t(lpvec.list[[t]]))
  
    my.eigen <- partial_eigen(adj.list[[t]], lp.dim, symmetric=TRUE)
    raw.scaled.list[[t]] <- my.eigen$vectors %*% abs(diag(my.eigen$values))^(1/2)
  
    my.candidate.svd <- svd(raw.scaled.list[[t]][1:lp.dim,])
    my.candidate.orthog <- my.candidate.svd$u %*% t(my.candidate.svd$v)
    svsd.orthog.list[[t]] <- my.candidate.orthog
  
    Xhat.list[[t]] <- raw.scaled.list[[t]] %*% t(svsd.orthog.list[[t]])
    
  }
```


### Collect simulation data (trajectories) across all nodes
```{r}

  trajs <- dynamic.node.num

  comp.types <- 4
  # number of different computations populating the data matrix

  data.table <- matrix(0, nrow=trajs, ncol=comp.types)
      
      for(i in 1:trajs){
      
        temp.mtx <- matrix(0, nrow=t.max, ncol=lp.dim)
        
        which.node <- i + lp.dim
        
        for(t in 1:t.max){
          temp.mtx[t,] <- Xhat.list[[t]][which.node,]
        }
      
###########################
      
          Y.mtx <- t(temp.mtx)[,2:t.max]
        
          Z.mtx <- rbind(rep(1,(t.max-1)),
                       t(temp.mtx)[,1:(t.max-1)])
                       
          Bhat.mtx <- Y.mtx %*% t(Z.mtx) %*% solve(Z.mtx %*% t(Z.mtx)); Bhat.mtx
        
          temp.decomp <- svd(Bhat.mtx[,2:3])
          temp.orthog <- temp.decomp$u %*% t(temp.decomp$v); temp.orthog
        
          target.orthog <- rot_mtx(hyper.thetas[[which.node]]); target.orthog
        
          data.table[i, 1] <- norm(Bhat.mtx[,2:3] - target.orthog, "F")/norm(target.orthog, "F")
          data.table[i, 2] <- norm(temp.orthog - target.orthog, "F")/norm(target.orthog, "F")
          data.table[i, 3] <- round(hyper.thetas[[which.node]], digits = 3)
          data.table[i, 4] <- round(norm(lpvec.list[[1]][which.node,],"2"), digits = 2)
        
      }
      data.table <- data.table[c(-dynamic.node.num,-dynamic.node.num/2),]
      # technicality: remove points for which we do not have at least two data points
      
      data.table <- as.data.frame(data.table)
      colnames(data.table) <- c("OLS_relFrob","OLS_ORTH_relFrob","hyp_Theta","lp_norm")
      
```

## Plot: rel. Frob. err. (points)
```{r}

  plot_rF_pts <- 
    ggplot(data.table, aes(x=OLS_relFrob, y=OLS_ORTH_relFrob)) +
            xlab("\nOLS rel. Frob. err.") + ylab("OLS ORTH rel. Frob. err.\n") +
            theme(legend.position="none", text = element_text(size=20)) +
            geom_point(inherit.aes = TRUE, size=1.5) 


  plot_rF_pts

```


## Plot: rel. Frob. err. (lp norm vals)
```{r}

    plot_rF_nums <-
          ggplot(data.table, aes(x=OLS_relFrob, y=OLS_ORTH_relFrob)) +
            xlab("\nOLS rel. Frob. err.") + ylab("OLS ORTH rel. Frob. err.\n") +
            theme(legend.position="none", text = element_text(size=20)) +
            geom_text(aes(label=lp_norm), check_overlap = TRUE, size=3.8)

    plot_rF_nums

```

# End of File

