---
title: "DyNet paper - Bitcoin Alpha data analysis"
author:
    name: Joshua Cape
    affiliation: University of Pittsburgh
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

Data synopsis: Bitcoin Alpha trust weighted signed (temporal) network

Data access: https://snap.stanford.edu/data/soc-sign-bitcoin-alpha.html

Data filename (local): bitcoinalpha.csv

# Load preamble
```{r message=FALSE}

  rm(list = ls())
  
  knitr::opts_chunk$set(comment = NA)
  
  source("code_dynet_preamble.R")
  
```


# Load data
```{r message=FALSE}

  soc_sign_bitcoin <-
    read_csv("soc-sign-bitcoinalpha.csv",
             col_names = c("source","target","rating","time"))

  my.data <- as.data.frame(soc_sign_bitcoin)
  my.data$time <- year(as_datetime(my.data$time))

  edge.table <- table(my.data$time); edge.table # counts number of links for each year
  my.years <- names(edge.table)[2:6] # specified years of interest: 2011, ..., 2015
  
```

```{r message=FALSE}

  # >>> Checks <<<
  #
      temp <- my.data[my.data$time %in% my.years,c("source","target")]; # restrict to specified years
      dim(as.data.frame(distinct(temp))) # number of distinct directed edges across time
      dim(as.data.frame(temp)) # number of total directed edges in aggregate graph
      rm(temp)
  
```

# Data preprocessing
```{r message=FALSE}

# To begin, keep largest connected component over all specified years
  my.g <- graph_from_edgelist(cbind(my.data$source, my.data$target), directed = TRUE)
  E(my.g)$weight <- my.data$rating
  
  g.components <- components(my.g, mode="weak")
  
  my.subgraph <- induced_subgraph(my.g, v=which(g.components$membership==1))
  # automatically includes E(my.subgraph)$weight
  
#   >>> Checks <<<
#
#   table(g.components$membership)  
#  
#   all(E(my.subgraph)$weight == my.data$rating[my.data$source %in% which(g.components$membership==1)])
#   induced subgraph automatically keeps corresponding edge weights
#    
#   table((my.data$source %in% which(g.components$membership==1)) ==
#         (my.data$target %in% which(g.components$membership==1)))
#   dim(my.data)[1]
#  
#   Keep my.data rows for which both source & target are in largest component (index)
#   which(g.components$membership==1)

```


## Investigate the aggregate graph
```{r message=FALSE}

  my.adj <- as_adjacency_matrix(my.subgraph, attr="weight")
  my.adj.sym <- (my.adj + t(my.adj))/2
  my.adj.sym.bin <- my.adj.sym
  my.adj.sym.bin[my.adj.sym > 0.001] <- 1
  my.adj.sym.bin[my.adj.sym < -0.001] <- -1
  my.adj.sym.bin[my.adj.sym > -0.001 & my.adj.sym < 0.001] <- 0

  temp <- graph_from_adjacency_matrix(my.adj.sym.bin, weighted = TRUE, mode="undirected")
  table(E(temp)$weight)
  is_connected(temp)
  
  my.g.agg <- induced_subgraph(temp, which(components(temp,"weak")$membership==1)); rm(temp)
  table(E(my.g.agg)$weight)
  is_connected(my.g.agg) # verifies connectivity of graph
  
  
  #set.seed(1234)
    ASE.agg <-
      embed_adjacency_matrix(my.g.agg, no=2, which="lm")$X        # compute raw ASE
    ASE.agg <- varimax(ASE.agg+1e-10, normalize=FALSE)$loadings   # align using varimax
    ASE.agg <- ASE.agg %*% diag(sign(ASE.agg[1,]))                # pos entries first components
    

```



```{r eval = FALSE, echo = FALSE}

# Investigate transformed adjacency matrices (of aggregate graph) 
  
  for(data.vals in c(my.adj, my.adj.sym, my.adj.sym.bin)){
    print(dim(data.vals))
    print(partial_eigen(data.vals, n=4)$values)
    print(svdr(data.vals, k=4)$d)
  }
  
# Print tables of edge weights (between -10 and 10)
    #table(as.numeric(as_adjacency_matrix(my.g, attr="weight")))
    #table(as.numeric(as_adjacency_matrix(my.subgraph, attr="weight")))
    #table(as.numeric(my.adj.sym))

```


# Process and individually embed the year graphs
Form graphs from all vertices and (weighted) edges appearing in a given, specified year.
Symmetrize the adjacency matrix, then binarize.
Keep the largest weakly connected component.
```{r message=FALSE}

set.seed(1234)

my.edge.data.list <- list()
my.year.graphs <- list()
    
  for(index in 1:length(my.years)){
    my.edge.data.list[[index]] <- my.data[my.data$time == my.years[index],]
    
    # Step 1
    my.year.graphs[[index]] <- graph_from_edgelist(cbind(my.edge.data.list[[index]]$source,
                                         my.edge.data.list[[index]]$target),
                                   directed = TRUE)
    E(my.year.graphs[[index]])$weight <- my.edge.data.list[[index]]$rating
    
    # Step 2
    temp1 <- components(my.year.graphs[[index]], mode="weak")
    temp2 <- which(temp1$membership ==
                     (which(table(temp1$membership) ==
                              max(table(temp1$membership)))))
    
    my.year.graphs[[index]] <- induced_subgraph(my.year.graphs[[index]],temp2)
    
  }
  
my.year.graphs.proc <- list()
my.year.ASE <- list()  
  
  for(index in 1:length(my.years)){
      my.adj <- as_adjacency_matrix(my.year.graphs[[index]], attr="weight")
      my.adj.sym <- (my.adj + t(my.adj))/2
    my.adj.sym.bin <- my.adj.sym
      my.adj.sym.bin[my.adj.sym > 0.001] <- 1
      my.adj.sym.bin[my.adj.sym < -0.001] <- -1
      my.adj.sym.bin[my.adj.sym > -0.001 & my.adj.sym < 0.001] <- 0
      
    my.year.graphs.proc[[index]] <-
      graph_from_adjacency_matrix(my.adj.sym.bin, weighted = TRUE, mode="undirected")
    
#    print(is_connected(my.year.graphs.proc[[index]])) # checks connectivity of graph
    
    # Update
    
        temp1 <- components(my.year.graphs.proc[[index]], mode="weak")
        temp2 <- which(temp1$membership ==
                         (which(table(temp1$membership) ==
                                  max(table(temp1$membership)))))
    my.year.graphs.proc[[index]] <- induced_subgraph(my.year.graphs.proc[[index]], temp2)

#    print(is_connected(my.year.graphs.proc[[index]])) # checks connectivity of graph

    my.year.ASE[[index]] <-
      embed_adjacency_matrix(my.year.graphs.proc[[index]], no=2, which="lm")$X  # compute raw ASE
    my.year.ASE[[index]] <-
      varimax(my.year.ASE[[index]]+1e-10, normalize=FALSE)$loadings             # align using varimax
    my.year.ASE[[index]] <-
      my.year.ASE[[index]] %*% diag(sign(my.year.ASE[[index]][1,]))             # + sign first components
  }

  my.year.graphs.proc[[6]] <- my.g.agg

```


```{r message=FALSE}

  mtx_print(
    sapply(my.year.graphs.proc, function(x){c(length(V(x)),length(E(x)))})
    )

mtx_print(rbind(
      format(signif(sapply(my.year.graphs.proc, function(x){table(E(x)$weight)})["-1",]/
        colSums(sapply(my.year.graphs.proc, function(x){table(E(x)$weight)})) * 100, 3))))

  mtx_print(
    rbind(
      format(round(sapply(my.year.graphs.proc, function(x){edge_density(x)}), 3), nsmall=3)
    )
  )

```

# All plots
```{r message=FALSE, warning=FALSE}

library(cowplot)

df.ASE.years <- list()
my.col.names <- c("x","y")

temp.vals <- 1
temp.sz <- 0.5

for(index in 1:length(my.years)){
  df.ASE.years[[index]] <- as.data.frame(my.year.ASE[[index]])
  colnames(df.ASE.years[[index]]) <- my.col.names
  
  assign(paste("viz.ASE.", index, sep=""), ggplot(df.ASE.years[[index]]) +
           geom_point(aes(x, y), size=temp.sz) + xlim(-temp.vals, temp.vals) + ylim(-temp.vals, temp.vals)
         )
}

  df.ASE.years.agg <- as.data.frame(ASE.agg)
  colnames(df.ASE.years.agg) <- my.col.names
  viz.ASE.agg <- ggplot(df.ASE.years.agg) + geom_point(aes(x, y), size=temp.sz) + xlim(-temp.vals, temp.vals) + ylim(-temp.vals, temp.vals)


my.plot.grid <-
  plot_grid(viz.ASE.1, viz.ASE.2, viz.ASE.3, viz.ASE.4, viz.ASE.5, viz.ASE.agg, labels="AUTO")

my.plot.grid

```



## End of File
